---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "800px",
  dpi=150
)

library(palmerpenguins)
library(ggplot2)
library(magick)
library(ggreveal)
library(patchwork)

out_width <- 9
out_height <- 5

p <-  ggplot(penguins[!is.na(penguins$sex),],
             aes(body_mass_g, bill_length_mm,
                 group=sex, color=sex)) +
      geom_point() +
      geom_smooth(method="lm", formula = 'y ~ x', linewidth=1) +
      facet_wrap(~species) +
      theme_minimal() +
      theme(plot.title = element_text(size = 16, hjust = 0.5))

p

plots_group <- reveal_save(reveal_groups(p + labs(title = "reveal_groups()")), 
                          "group.png", width=out_width, height=out_height, 
                          bg = "white", path = tempdir(), dpi=150)
plots_panel <- reveal_save(reveal_panels(p + labs(title = "reveal_panels()")), 
                          "panel.png", width=out_width, height=out_height, 
                          bg = "white", path = tempdir(), dpi=150)
plots_layer <- reveal_save(reveal_layers(p + labs(title = "reveal_layers()")), 
                          "layer.png", width=out_width, height=out_height, 
                          bg = "white", path = tempdir(), dpi=150)


p2 <- ggplot(penguins[!is.na(penguins$sex),],
             aes(x = year,
                 group=species, 
                 fill=species)) +
      geom_bar(position = position_dodge(.9)) +
      theme_minimal() +
      theme(plot.title = element_text(size = 16, hjust = 0.5)) +
      labs(title = "reveal_x()")

plots_x <- reveal_save(reveal_x(p2), 
                          "x.png", width=out_width, height=out_height,
                          bg = "white", path = tempdir(), dpi=150)

# Patchwork
pw1 <- ggplot(penguins[!is.na(penguins$sex),],
             aes(body_mass_g, bill_length_mm,
                 group=species,
                 color=species)) +
      geom_point() +
      guides(color = "none") +
      theme_minimal()


pw2 <- ggplot(penguins[!is.na(penguins$sex),],
             aes(x = year,
                 group=species, 
                 fill = species)) +
      geom_bar(position = position_dodge(.9)) +
      theme_minimal() 

pw3 <- ggplot(penguins[!is.na(penguins$sex),],
             aes(flipper_length_mm ,
                 fill=species)) +
      geom_histogram(position = "identity") +
      theme_minimal() 

pw <- pw1 + 
  (pw2/pw3) + 
  plot_layout(guides = "collect") +
  plot_annotation(title = "reveal_patchwork()",
                  tag_levels = "a",
                  theme = theme(plot.title = element_text(size = 16, hjust = 0.5))) 
  
plots_pw <- reveal_save(reveal_patchwork(pw), 
                          "pw.png", width=out_width, height=out_height, 
                          bg = "white", path = tempdir(), dpi=150)
                     

img_list <- lapply(c(plots_group, plots_panel, plots_layer, plots_x, plots_pw), image_read)
img_joined <- image_join(img_list)
gif <- image_animate(img_joined, fps = 1)
```


# ggreveal 

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/ggreveal)](https://cran.r-project.org/package=ggreveal)
[![Downloads](https://cranlogs.r-pkg.org/badges/grand-total/ggreveal)](https://cran.r-project.org/package=ggreveal)
[![R-CMD-check](https://github.com/weverthonmachado/ggreveal/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/weverthonmachado/ggreveal/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`ggreveal` is an R package that makes it easy to present data on ggplot2 graphs incrementally. The functions in this package take a plot and break it down into a series of intermediary plots that can be shown in sequence (e.g. in different slides). Like this:

```{r, message=FALSE, warning=FALSE, echo=FALSE}
gif
```

*Why* would you want to do that? Because it's fun and very useful for teaching and giving talks. 


## Installation

``` r
install.packages("ggreveal")
```

## Usage

Create a graph with ggplot2 as usual.

```r
library(palmerpenguins)
library(ggplot2)
library(ggreveal)

p <-  ggplot(penguins[!is.na(penguins$sex),],
             aes(body_mass_g, bill_length_mm,
                 group=sex, color=sex)) +
      geom_point() +
      geom_smooth(method="lm", formula = 'y ~ x', linewidth=1) +
      facet_wrap(~species) +
      theme_minimal()
p
```
```{r, message=FALSE, warning=FALSE, echo=FALSE}
p
```

Then use one of the `reveal_*` functions (`reveal_layers()`, `reveal_aes()`, `reveal_groups()`, `reveal_panels()`, `reveal_x()`) to obtain a list of plots that show elements incrementally. 

```{r, out.width="70%"}
plot_list <- reveal_groups(p)
plot_list
```

You can also use `reveal_patchwork()` to incrementally reveal the individual plots in a patchwork object.

The most common use case is to show the plots in different slides of a presentation. You can work directly with the list of plots returned by the functions (useful for Quarto and R Markdown presentations), or save the plots as separate files and include them in your presentation software of choice. `reveal_save()` is a convenient function to save the plots in a list:

```r
reveal_save(plot_list, "myplot.png", width = 8, height = 4)
```
```
── Saving incremental plots ──

✔ myplot_0.png
✔ myplot_1.png
✔ myplot_2_last.png
```


# <a id="why"></a> Wait, can't ggplot2 do this already?

Yes and no. `ggplot2` is composable by design, so it is straightforward to do some incremental reveal out of the box. For example, you can create the plot in steps and save the steps separately:

```r
data("mtcars")
p1 <- ggplot(mtcars, 
            aes(mpg, wt)) +
     geom_point() 
     
p2 <- p1 + geom_smooth(method="lm", formula="y~x") 
```

The problem is that, as you add layers and other elements, several visual aspects of the plot — e.g. range of axes, legends — can, and will often, change. Showing how the layout changes as you add elements is useful for teaching *how to make graphs in ggplot2*, but it is distracting when you want to focus on the *results*. Keeping the layout fixed is sometimes easy (e.g. setting the `limits` argument for the x/y scale), but in many cases it requires tinkering with the internals of the ggplot object. 



