---
title: "Get Started"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Get Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  dpi = 200
)
library(magick)
library(purrr)

create_gif <- function(plot_list, width=9, height=5.1) {
  temp_files <- tempfile(fileext = paste0("_", seq_along(plot_list), ".png"))
  plot_list <- plot_list  |>  
                imap(~{
                .x + ggtitle(paste0("plot_list[[", .y, "]]"))
                })
  for (i in seq_along(plot_list)) {
    ggsave(temp_files[i], plot_list[[i]], width = width, height = height, dpi = 200)
  }

  imgs <- image_read(temp_files)
  image_animate(imgs, fps = 1)
}
```

```{r setup}
library(ggreveal)
```


The goal of ggreveal is to make it easy to incrementally show parts of a ggplot. The package offers different ways to split a finished plot into steps. 

# Examples

Let's create a plot that to illustrate different ways to reveal elements. We will use the `penguins` dataset from the `palmerpenguins` package, which contains measurements of penguins from three different species.

```{r, message=FALSE, warning=FALSE, fig.width=9, fig.height=5}
library(palmerpenguins)
library(ggplot2)

penguins <-penguins[!is.na(penguins$sex),]

p1 <-  ggplot(penguins,
             aes(x = body_mass_g, 
                 y = bill_length_mm,
                 color = sex)) +
      geom_point() +
      geom_smooth(method="lm", formula = 'y ~ x', linewidth=1) +
      facet_wrap(~species) +
      theme_minimal() 

p1
```

We have a scatter plot of body mass vs bill length, colored by sex and faceted by species. On top of the points, we have linear regression lines for each combination of sex and species.

## Reveal by aesthetic

Let's say that during a presentation we want to highlight the differences between female and male pennguins. The `sex` variable is mapped to the `color` aesthetic, so we can use it with `reveal_aes()`. 

```{r}
plot_list <- reveal_aes(p1, aes = "color") 
```

`plot_list` is a list of 3 plots: the first shows only layout elements (axes, legend, etc) but no data; the second adds the data for female penguins; and the third adds the data for male penguins, thus arriving at the original plot.

```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot_list |> 
  create_gif()
```

*Note: ggreveal does not produce animations. The functions return a list of plots, which are shown here in animated gifs for sake of concision.*

The `reveal_aes()` function is quite flexible and can be used to reveal pretty much any aesthetic mapping. Suppose we map sex the *shape* aesthetic instead: [^fn1]

[^fn1]: Even though geom_smooth() does not use the shape aesthetic, `reveal_aes()` correctly groups the lines by sex because the shape aesthetic was defined in the global aes(). If we had instead defined shape inside geom_point(), `reveal_aes()` would first reveal the points for each sex, then all the regression lines together. Try it!

```r
p2 <-  ggplot(penguins,
             aes(x = body_mass_g, 
                 y = bill_length_mm,
                 shape= sex)) +
      geom_point() +
      geom_smooth(method="lm", formula = 'y ~ x', linewidth=1) +
      scale_shape_manual(values = c(1,15))
      facet_wrap(~species) +
      theme_minimal() 

plot_list <- reveal_aes(p2, aes = "shape")
plot_list
```
```{r, message=FALSE, warning=FALSE, echo=FALSE}
p2 <-  ggplot(penguins,
             aes(x = body_mass_g, 
                 y = bill_length_mm,
                 shape = sex)) +
      geom_point() +
      geom_smooth(method="lm", formula = 'y ~ x', linewidth=1) +
      facet_wrap(~species) +
      scale_shape_manual(values = c(1,15)) +
      theme_minimal() 

reveal_aes(p2, aes = "shape") |> 
  create_gif()
```

Or say we want to reveal categories in the x axis:

```r
p3 <-  ggplot(penguins,
             aes(x = sex, y = bill_length_mm )) +
      geom_boxplot() +
      theme_minimal() 

plot_list <- reveal_aes(p3, aes = "x")
plot_list
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, out.width = "55%"}
p3 <-  ggplot(penguins,
             aes(x = sex, y = bill_length_mm )) +
      geom_boxplot() +
      theme_minimal() 

reveal_aes(p3, aes = "x") |> 
  create_gif(5,5)
```

`reveal_x()` is a shortcut for `reveal_aes(aes = "x")`, and similarly for `reveal_y()`. Revealing axes works better when the x or y variable is discrete, or has few values, otherwise you can get a lot of intermediary plots. If you find yourself needing to incrementally reveal a large amount of values (e.g. a time series developing over the x axis), you might want to use [`gganimate`](https://gganimate.com/) instead. 

There is also reveal_groups


## Reveal by panel/facet

Returning to our first plot, we might instead want to draw attention to differences across species, and thus reveal each panel (facet) at a time with `reveal_panels()`. Now we get four plots: one "blank", then one plot for each of the three species revealed cumulatively. Again, the last element in the list corresponds to the original plot:

```r
plot_list <- reveal_panels(p1)
plot_list
```
```{r, message=FALSE, warning=FALSE, echo=FALSE}
reveal_panels(p1) |> 
  create_gif()
```

## Reveal by layer

Finally, we might want to first show the raw data and then add the regression lines. These elements are defined by different layers, so we can use `reveal_layers()`:

```r
plot_list <- reveal_layers(p1)
plot_list
```
```{r, message=FALSE, warning=FALSE, echo=FALSE}
reveal_layers(p1) |> 
  create_gif()
```

## Reveal a patchwork object

Inetad of element sin a single plot, we can also reveal elements in a patchwork object. For example, we can combine our first plot with a bar plot of the number of penguins by species:

```{r, message=FALSE, warning=FALSE}
library(patchwork)  

p4 <- ggplot(penguins, 
             aes(x = species)) +
      geom_bar() +
      theme_minimal()

pw <- p1 / (p3 + p4)

#reveal_patchwork(pw) |> 
  #create_gif(9,8)
```

# Rearranging the order of elements

